### 2024-12-12 2024-5-13/sections/heuristic.tex
first finished edition

\item We estimate the sizes of the two CEFAs $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ constructed from $\ufld_1(e), \ufld_2(e)$ respectively as pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$, where $\snum_1(e)$ and $\rnum_1(e)$ are the estimates of the number of states and registers in $\aut_{\ufld_1(e)}$, similarly for $\snum_2(e)$ and $\rnum_2(e)$. 
%
\item From the construction in Section~\ref{sec:algorithm}, 
two NFAs $\cU_{\aut_{\ufld_2(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are constructed from $\aut_{\ufld_1(e)}$ and $\aut_{\ufld_2(e)}$ by dropping the accepting conditions and ignoring the characters. As a result, the alphabets of $\cU_{\aut_{\ufld_2(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are the set of integer vectors  $\myvec{v}$ occurring in the transitions of $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ respectively. The integer vectors of  $\aut_{\ufld_1(e)}$ are of form $(\underbrace{0, \cdots, 0}_{i-1}, 1, \underbrace{0, \cdots, 0}_{r'-i-1})$, where $i \in [r']$. Consequently, the size of the alphabet of $\cU_{\aut_{\ufld_1(e)}}$ is bounded by $\rnum_1(e)$. Similarly for $\cU_{\aut_{\ufld_1(e)}}$. 
Finally, we estimate the sizes (number of transitions) of the NFAs $\cU_{\aut_{\ufld_1(e)}}, \cU_{\aut_{\ufld_2(e)}}$ as $sz_1 = (\snum_1(e))^2 \cdot \rnum_1(e)$ and $sz_2 = (\snum_2(e))^2 \cdot \rnum_2(e)$ respectively.
%
\item If $sz_1 \le sz_2$, then we choose to unfold the counting operators in $e_1$. Otherwise, we choose to unfold the counting operator $^{\{m,n\}}$. 
\end{itemize}

@@

It remains to show how to compute the pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$ from $e=(e_1)^{\{m,n\}}$. Let $e=(e_1)^{\{m,n\}}$ be a regular expression where $e_1$ is a register-representable regular expression. Then we compute $\snum'_1(e_1)$ and $(\snum'_2(e_1), \rnum'_2(e_1))$ by a structural induction on $e_1$. Finally, let $(\snum_1(e), \rnum_1(e)) = (\snum'_1(e_1), 1)$ and $(\snum_2(e), \rnum_2(e)) = (\snum'_2(e_1), \rnum'_2(e_1))$.
\begin{itemize}
  \item If $e_1 = \emptyset$ or $e_1 = \epsilon$, then $\snum'_1(e_1)  = 0$ and $(\snum'_2(e_1), \rnum'_2(e_1))  = (0, 0)$. 
%
 %
  \item If $e_1 = e_2 \concat e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = e_2 + e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3)+1)).
 \end{array}
  $$
 (Recall that the construction of $\aut_1 \cup \aut_2$ introduces one additional register.)
%
 \item If $e_1 = e_2 \cap e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) \times \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) \times \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = \overline{e_2}$, then $\snum'_1(e_1) = 2^{\snum'_1(e_2)}$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \cdot 2^{\snum'_2(e_2)}, 0).
% \end{array}
  $
 (The exponential blow-up is explained by the fact that the determinization is needed for the complementation.) 
%
%  \item If $\regex = \overline{\regex_1}$, then $s_\regex = 2^{s_{\regex_1}}$ (since the determinization is needed for the complement) and $r_\regex = 0$.
%
 \item If $e_1 = (e_2)^*$, then $\snum'_1(e_1) = \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), 0).
% \end{array}
  $
%  \item If $\regex = \regex_1^*$, then $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 0$.
%  \item $\regex = \regex_1 \backslash \regex_2$ can be transferred to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the size can be computed by the formula of $\cap$ and $\bar{}$.
 \item If $e_1 = (e_2)^{\{m', n'\}}$ or $e_1 = (e_2)^{\{n', \infty\}}$, then $\snum'_1(e_1) = n' \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), n).
% \end{array}
  $
%  \item For $\regex = \regex_1^{\{m, n\}}$ or $\regex = \regex_1^{\{n,\infty\}}$, $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 1$.
\end{itemize}

@@

Without loss of generality, we can assume that there are no transitions out of $F$ in $\aut_{e_1}$. If this is not the case, then a new state $q_f$ can be introduced to transform $\aut_{e_1}$ into a CEFA satisfying this constraint, where $q_f$ is the unique accepting state. 
Then we define $\cB= (R \cup \{r'\}, Q \cup \{q_s\}, \Sigma, \delta', I, Q \cup \{q_s\}, \alpha')$, where
\begin{itemize}
  \item $r'$ is a new register not in $R$, 
  \item $q_s$ is a new state not in $Q$,
  \item $\delta'$ comprises the following transitions, 
  \begin{itemize}
    \item $(q, a, q', (\vec{v}, 0))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \not \in F$,
%
    \item $(q, a, q_s,(\vec{0}, 0))$ such that there do \emph{not} exist $a$-labeled transitions out of $q$, that is, for every $q'$ and $\vec{v}$, we have $(q, a, q', \vec{v}) \not\in \delta$,
%
    \item $(q, a, q', (\vec{v}, 1))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \in F$,
%
    \item $(q_s, a, q_s, (\vec{0}, 0))$ such that $a \in \Sigma$, 
  \end{itemize}
  \item $\alpha'$ is defined as $r' = 1 \rightarrow \neg \alpha$.
\end{itemize}
Intuitively, $q_s$ is a sink state that deals with the situations where there are no runs, and $r'$ is a register newly introduced so that the different cases of the non-accepting runs of $\aut_{e_1}$ can be captured by a single accepting condition $\alpha'$.

@@

@@

Recall that in Section~\ref{sec:algorithm}, the satisfiability of the original RECL constraint is reduced to the satisfiability of the existential LIA formula
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
Let us use $\theta$ to denote this formula. 
Then we utilize some SMT solver, say Z3, to solve the satisfiability of $\theta$. If Z3 reports ``sat'', then it also returns a model, namely, an assignment $\amodel$ of natural numbers to the free integer variables in $\theta$. Then we utilize $\amodel$ to generate a model $\amodel'$ of the original RECL constraint $\varphi$, i.e., an assignment of strings to the string variables so that $\varphi$ is evaluated to true under $\amodel'$.

Note that for each $i \in [n]$, the free variables $\fZ_{\costset_{\cC_i}}$ in $\theta$ correspond to the number of occurrences of characters in $\costset_{\cC_i}$, i.e. the alphabet of the NFA $\cC_i$. Moreover, recall that in $\cC_i$, the characters are just the integer vectors and the characters in the original CEFAs $\cB_i$ used to construct $\cC_i$ are ignored. 
When we utilize $\amodel$ and the CEFAs $\cB_i$ to generate $\amodel'$, for each $i \in [n]$, we record the remaining occurrences of vectors $\myvec{v}$ in the transitions of the CEFAs $\cB_i$ as a number $\nvec_{i, \myvec{v}}$ and
each time when a transition $(q, a, q', \myvec{v})$ in $\cB_i$ is traversed, $\nvec_{i, \myvec{v}}$ is decremented. Note that initially $\nvec_{i, \myvec{v}}$ is set to $\amodel(\anivarz_{i, \myvec{v}})$. The searching process terminates when $\nvec_{i, \myvec{v}}$ becomes $0$ for every $\myvec{v}$. We record the characters in the traversed transitions and obtain a string for the string variable $x_i$ when the searching process terminates. As a result, we obtain a model of $\varphi$ in the end.

It turns out that this model generation process is inefficient, mainly due to the fact that the information about the occurrences of vectors $\myvec{v}$ in $\cB_i$ in the searching process is too coarse. In the sequel, we would like to refine this process by using the LIA formulas to obtain the information about the number of occurrences of transitions, instead of just the occurrences of integer vectors, and utilize this information to guide the search process. 
Specifically, for each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, we introduce an integer variable $\anivarz_{i, (q, a, q', \myvec{v})}$ to denote the number of occurrences of the transition, and solve the satisfiability of the following formula,  
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right]\ \wedge \\
\ \ \bigwedge \limits_{i \in [n]} \bigwedge \limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} = \sum \limits_{(q, a, q', \myvec{v}) \mbox{ occurs in } \cB_i} \anivarz_{i, (q, a, q', \myvec{v})}.
\end{array}
\]
If an SMT solver reports that the aforementioned formula is satisfiable, then the solver can also return a model, which assigns to each variable $\anivarz_{i, (q, a, q', \myvec{v})}$ a natural number. 
%For each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, let us use $nt_{i, (q, a, q', \myvec{v})}$ to denote the value of $\anivarz_{i, (q, a, q', \myvec{v})}$ in the model.
Then we utilize these values to generate a model (an assignment of strings to string variables) of $\varphi$. 
Let $i \in [n]$. We use a function $\lambda_i$ to record the number of occurrences of the transitions in $\cB_i$. During the search process for $\cB_i$, each time when a transition $(q, a, q', \myvec{v})$ is traversed, $\lambda_i$ is updated by decrementing $\lambda_i((q, a, q', \myvec{v}))$, moreover, $a$ is also recorded. Initially, for each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, $\lambda_i((q, a, q', \myvec{v}))$ is set to be the value of $\anivarz_{i, (q, a, q', \myvec{v})}$ in the model. When a final state of $\cB_i$ is reached and $\lambda_i((q, a, q', \myvec{v}))$ becomes zero for each transition $(q, a, q', \myvec{v})$ in $\cB_i$, then a desired string is generated for the string variable $x_i$.

>>>>

\item We estimate the sizes of the two CEFAs $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ constructed from $\ufld_1(e), \ufld_2(e)$ respectively as pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$, where $\snum_1(e)$ and $\rnum_1(e)$ are the estimates of the number of states and registers in $\aut_{\ufld_1(e)}$, similarly for $\snum_2(e)$ and $\rnum_2(e)$. 
%
\item From the construction in Section~\ref{sec:algorithm}, 
two NFAs $\cU_{\aut_{\ufld_\hdh{1}(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are constructed from $\aut_{\ufld_1(e)}$ and $\aut_{\ufld_2(e)}$ by dropping the accepting conditions and ignoring the characters. As a result, the alphabets of $\cU_{\aut_{\ufld_\hdh{1}(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are the set of integer vectors \sout{$\myvec{v}$} occurring in the transitions of $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ respectively. \sout{The integer vectors of  $\aut_{\ufld_1(e)}$ are of form $(\underbrace{0, \cdots, 0}_{i-1}, 1, \underbrace{0, \cdots, 0}_{r'-i-1})$, where $i \in [r']$.} \hdh{These integer vectors are of form $(0,\cdots,0,1,0,\cdots, 0)$ whose length is the number of registers.} Consequently, the size of the alphabet of $\cU_{\aut_{\ufld_1(e)}}$ is bounded by $\rnum_1(e)$. Similarly for $\cU_{\aut_{\ufld_\hdh{2}(e)}}$. 
Finally, we estimate the sizes (number of transitions) of the NFAs $\cU_{\aut_{\ufld_1(e)}}, \cU_{\aut_{\ufld_2(e)}}$ as $sz_1 = (\snum_1(e))^2 \cdot \rnum_1(e)$ and $sz_2 = (\snum_2(e))^2 \cdot \rnum_2(e)$ respectively.
%
\item If $sz_1 \le sz_2$, then we choose to unfold the counting operators in $e_1$. Otherwise, we choose to unfold the counting operator $^{\{m,n\}}$. 
\end{itemize}

@@

It remains to show how to compute the pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$ from $e=(e_1)^{\{m,n\}}$. Let $e=(e_1)^{\{m,n\}}$ be a regular expression where $e_1$ is a register-representable regular expression. Then we compute $\snum'_1(e_1)$ and $(\snum'_2(e_1), \rnum'_2(e_1))$\hdh{(TODO:Change all the $num'$ to $num$, to avoid ambiguity)} by a structural induction on $e_1$, \hdh{in condition, $\rnum_1(e_1)$ is set to $1$.} \sout{Finally, let $(\snum_1(e), \rnum_1(e)) = (\snum'_1(e_1), 1)$ and $(\snum_2(e), \rnum_2(e)) = (\snum'_2(e_1), \rnum'_2(e_1))$.}
\begin{itemize}
  \item If $e_1 = \emptyset$ or $e_1 = \epsilon$, then $\snum'_1(e_1)  = 0$ and $(\snum'_2(e_1), \rnum'_2(e_1))  = (0, 0)$. 
%
  \item \hdh{If $e_1 = a$, then $\snum'_1(e_1) = 1$ and ($\snum'_2(e_1), \rnum'_2(e_1)) = (n, 0)$.}
 %
  \item If $e_1 = e_2 \concat e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = e_2 + e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3)+1)).
 \end{array}
  $$
 (Recall that the construction of $\aut_1 \cup \aut_2$ introduces one additional register.)
%
 \item If $e_1 = e_2 \cap e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) \times \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) \times \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = \overline{e_2}$, then $\snum'_1(e_1) = 2^{\snum'_1(e_2)}$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \cdot 2^{\snum'_2(e_2)}, 0).
% \end{array}
  $
 (The exponential blow-up is explained by the fact that the determinization is needed for the complementation.) 
%
%  \item If $\regex = \overline{\regex_1}$, then $s_\regex = 2^{s_{\regex_1}}$ (since the determinization is needed for the complement) and $r_\regex = 0$.
%
 \item If $e_1 = (e_2)^*$, then $\snum'_1(e_1) = \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), 0).
% \end{array}
  $
%  \item If $\regex = \regex_1^*$, then $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 0$.
%  \item $\regex = \regex_1 \backslash \regex_2$ can be transferred to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the size can be computed by the formula of $\cap$ and $\bar{}$.
 \item If $e_1 = (e_2)^{\{m', n'\}}$ or $e_1 = (e_2)^{\{n', \infty\}}$, then $\snum'_1(e_1) = n' \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), n).
% \end{array}
  $
%  \item For $\regex = \regex_1^{\{m, n\}}$ or $\regex = \regex_1^{\{n,\infty\}}$, $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 1$.
\end{itemize}

@@

Without loss of generality, we can assume that there are no transitions out of $F$ in $\aut_{e_1}$. If this is not the case, then a new state $q_f$ can be introduced to transform $\aut_{e_1}$ into a CEFA satisfying this constraint, where $q_f$ is the unique accepting state.\footnote{\hdh{Note that if any initial states are in $F$, a new state $q_I$ without transitions out can be introduced and add to $F$, to accept empty string. Then we can remove all original initial states out of $F$.}} 
Then we define $\cB= (R \cup \{r'\}, Q \cup \{q_s\}, \Sigma, \delta', I, Q \cup \{q_s\}, \alpha')$, where
\begin{itemize}
  \item $r'$ is a new register not in $R$, 
  \item $q_s$ is a new state not in $Q$,
  \item $\delta'$ comprises the following transitions, 
  \begin{itemize}
    \item $(q, a, q', (\vec{v}, 0))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \not \in F$,
%
    \item $(q, a, q_s,(\vec{0}, 0))$ such that there do \emph{not} exist $a$-labeled transitions out of $q$, that is, for every $q'$ and $\vec{v}$, we have $(q, a, q', \vec{v}) \not\in \delta$,
%
    \item $(q, a, q', (\vec{v}, 1))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \in F$,
%
    \item $(q_s, a, q_s, (\vec{0}, 0))$ such that $a \in \Sigma$, 
  \end{itemize}
  \item $\alpha'$ is defined as $r' = 1 \rightarrow \neg \alpha$.
\end{itemize}
Intuitively, $q_s$ is a sink state that deals with the situations where there are no runs, and $r'$ is a register newly introduced so that the different cases of the non-accepting runs of $\aut_{e_1}$ can be captured by a single accepting condition $\alpha'$.

@@

\hdh{\paragraph*{Combination of under- and over-approximation}
Since these two types of approximation influence each other, we separate them apart and run them parallelly.
}

@@

\hdh{The naive approach of finding satisfied string models is to search the accepted words on CEFAs based on the integer models of vectors, which causes an exponential blow-up of searching space. Specifically, } recall that in Section~\ref{sec:algorithm}, the satisfiability of the original RECL constraint is reduced to the satisfiability of the existential LIA formula
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
Let us use $\theta$ to denote this formula. 
Then we utilize some SMT solver, say Z3, to solve the satisfiability of $\theta$. If Z3 reports ``sat'', then it also returns a model, namely, an assignment $\amodel$ of natural numbers to the free integer variables in $\theta$. Then we utilize $\amodel$ to generate a model $\amodel'$ of the original RECL constraint $\varphi$, i.e., an assignment of strings to the string variables so that $\varphi$ is evaluated to true under $\amodel'$.

Note that for each $i \in [n]$, the free variables $\fZ_{\costset_{\cC_i}}$ in $\theta$ correspond to the number of occurrences of characters in $\costset_{\cC_i}$, i.e. the alphabet of the NFA $\cC_i$. Moreover, recall that in $\cC_i$, the characters are just the integer vectors and the characters in the original CEFAs $\cB_i$ used to construct $\cC_i$ are ignored. 
When we utilize $\amodel$ and the CEFAs $\cB_i$ to generate $\amodel'$, for each $i \in [n]$, we record the remaining occurrences of vectors $\myvec{v}$ in the transitions of the CEFAs $\cB_i$ as a number $\nvec_{i, \myvec{v}}$ and
each time when a transition $(q, a, q', \myvec{v})$ in $\cB_i$ is traversed, $\nvec_{i, \myvec{v}}$ is decremented. Note that initially $\nvec_{i, \myvec{v}}$ is set to $\amodel(\anivarz_{i, \myvec{v}})$. The searching process terminates when $\nvec_{i, \myvec{v}}$ becomes $0$ for every $\myvec{v}$. We record the characters in the traversed transitions and obtain a string for the string variable $x_i$ when the searching process terminates. As a result, we obtain a model of $\varphi$ in the end. \hdh{In the worst case, we need search O($\prod_{i=0}^{n} \amodel(\anivarz_{i, \myvec{v}})$) times to find an accepted word.}

It turns out that this model generation process is inefficient, mainly due to the fact that the information about the occurrences of vectors $\myvec{v}$ in $\cB_i$ in the searching process is too coarse. In the sequel, we would like to refine this process by using \sout{the LIA formulas to obtain the information about} the number of occurrences of transitions, instead of just the occurrences of integer vectors, \sout{and utilize this information} to guide the search process. 
Specifically, for each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, we introduce an integer variable $\anivarz_{i, (q, a, q', \myvec{v})}$ to denote the number of occurrences of the transition, and solve the satisfiability of the following formula,  
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right]\ \wedge \\
\ \ \bigwedge \limits_{i \in [n]} \bigwedge \limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} = \sum \limits_{(q, a, q', \myvec{v}) \mbox{ occurs in } \cB_i} \anivarz_{i, (q, a, q', \myvec{v})}.
\end{array}
\]
If an SMT solver reports that the aforementioned formula is satisfiable, then the solver can also return a model, which assigns to each variable $\anivarz_{i, (q, a, q', \myvec{v})}$ a natural number. 
%For each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, let us use $nt_{i, (q, a, q', \myvec{v})}$ to denote the value of $\anivarz_{i, (q, a, q', \myvec{v})}$ in the model.
Then we utilize these values to generate a model (an assignment of strings to string variables) of $\varphi$. 
Let $i \in [n]$. We use a function $\lambda_i$ to record the number of occurrences of the transitions in $\cB_i$. During the search process for $\cB_i$, each time when a transition $(q, a, q', \myvec{v})$ is traversed, $\lambda_i$ is updated by decrementing $\lambda_i((q, a, q', \myvec{v}))$, moreover, $a$ is also recorded. Initially, for each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, $\lambda_i((q, a, q', \myvec{v}))$ is set to be the value of $\anivarz_{i, (q, a, q', \myvec{v})}$ in the model. When a final state of $\cB_i$ is reached and $\lambda_i((q, a, q', \myvec{v}))$ becomes zero for each transition $(q, a, q', \myvec{v})$ in $\cB_i$, then a desired string is generated for the string variable $x_i$.

### 2024-12-12 2024-5-13/sections/implementation.tex
first finished edition

\medskip
\noindent
\emph{NestC benchmark suite.} 
We generate this benchmark suite by utilizing the string fuzzing tool \textsc{stringfuzz}~\cite{stringfuzz}. Note that \textsc{stringfuzz} is slightly extended by adding the counting operators.
%This benchmark suite is generated by string fuzzing test tool \textsc{stringfuzz} \cite{stringfuzz} extended with the counting operator. 
For the benchmarks that contain the nestings of counting operators, we use the following template
\[x\in e_{nest}\wedge x\in e_{sani}\wedge |x|>50\]
where $e_{nest}$ is of the form 
\[(e_1^{\{m_1, n_1\}}e_2^{\{m_2, n_2\}}e_3^{\{m_3, n_3\}}e_4^{\{m_4, n_4\}})^{\{m_5, n_5\}},\]
such that $0 \leq m_i\leq n_i\leq 20$ for each $i \in [5]$ and $e_i$ contains no counting operators for each $i  \in [4]$,
moreover, 
\[e_{sani} \equiv \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}.\]
Based on this template, we generate 500 instances.

For the benchmarks that contain the nestings of counting operators and a complement operator, we utilize the following template
\[x \in e_{comp1} \wedge x \in e_{comp2} \wedge x \in e_{sani} \wedge |x| > 50, \]
where both $e_{comp1}$ and $e_{comp2}$ are of the form $\overline{e_1^{\{m_1,n_1\}}e_2^{\{m_2,n_2\}}e_3^{\{m_3,n_3\}} e_4^{\{m_4, n_4\}}}$, such that for each $i \in [4]$, $20 \leq m_i\leq n_i\leq 1000$ and $e_i$ contains no counting operators. Note that two regular expressions $e_{comp1} $ and $e_{comp2}$ are introduced to have a better balance between the numbers of satisfiable and unsatisfiable instances. Based on this template, we also generate 500 instances.

In total, the NestC benchmark suite contains 1000 instances.

@@

\begin{table}[htbp]
  \import{tables}{table_heuristics_model_gen.tex}
  \caption{Evaluation of the optimizations for the model generation in Section~\ref{sec-opt-sol-gen}}\label{tab:results_model-gen}
\end{table}

>>>>

\medskip
\noindent
\emph{NestC benchmark suite.} 
We generate this benchmark suite by utilizing the string fuzzing tool \textsc{stringfuzz}~\cite{stringfuzz}. Note that \textsc{stringfuzz} is slightly extended by adding the counting operators.
%This benchmark suite is generated by string fuzzing test tool \textsc{stringfuzz} \cite{stringfuzz} extended with the counting operator. 
For the benchmarks that contain the nestings of counting operators, we use the following template
\[x\in e_{nest}\wedge x\in e_{sani}\wedge |x|>50\]
where $e_{nest}$ is of the form 
\[(e_1^{\{m_1, n_1\}}e_2^{\{m_2, n_2\}}e_3^{\{m_3, n_3\}}e_4^{\{m_4, n_4\}})^{\{m_5, n_5\}},\]
such that $0 \leq m_i\leq n_i\leq 20$ \hdh{is chosen randomly} for each $i \in [5]$ and $e_i$ contains no counting operators for each $i  \in [4]$,
moreover, 
\[e_{sani} \equiv \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}.\]
Based on this template, we generate 500 instances.

For the benchmarks that contain the nestings of counting operators and a complement operator, we utilize the following template
\[x \in e_{comp1} \wedge x \in e_{comp2} \wedge x \in e_{sani} \wedge |x| > 50, \]
where both $e_{comp1}$ and $e_{comp2}$ are of the form $\overline{e_1^{\{m_1,n_1\}}e_2^{\{m_2,n_2\}}e_3^{\{m_3,n_3\}} e_4^{\{m_4, n_4\}}}$, such that for each $i \in [4]$, $20 \leq m_i\leq n_i\leq 1000$ \hdh{is chosen randomly} and $e_i$ contains no counting operators. Note that two regular expressions $e_{comp1} $ and $e_{comp2}$ are introduced to have a better balance between the numbers of satisfiable and unsatisfiable instances. Based on this template, we also generate 500 instances.

In total, the NestC benchmark suite contains 1000 \hdh{tough} instances \hdh{which may not be practical in real-world programs}.

@@

\begin{table}[htbp]
  \centering
  \import{tables}{table_heuristics_model_gen.tex}
  \caption{Evaluation of the optimizations for the model generation in Section~\ref{sec-opt-sol-gen}}\label{tab:results_model-gen}
\end{table}
